---
title: ADR Activities
subtitle: Hikes and Adventure Sports
summary: Hikes and Adventure Sports
author: Harshvardhan
date: '2022-04-20'
slug: my-activities
categories:
  - life
tags: []
---

# Map of Cities Touched

```{r, include = FALSE}
library(leaflet)
library(tidyverse)
places = tribble(
  ~name, ~lat, ~lng,
  "Jhumri Tilaiya", 24.4315, 85.5310,
  "Indore", 22.7196, 75.8577,
  "Ranchi", 23.3441, 85.3096,
  "Dhanbad", 23.7957, 86.4304,
  "Kolkata", 22.5726, 88.3639,
  "New Delhi", 28.6139, 77.2090,
  "Amritsar", 31.6340, 74.8723,
  "Puri", 19.8135, 85.8312,
  "Goa", 15.4909, 73.8278,
  "Gokarna", 14.5479, 74.3188,
  "Mumbai", 19.0760, 72.8777,
  "Udaipur", 24.5854, 73.7125,
  "Ujjain", 23.1765, 75.7885,
  "Agra", 27.1767, 78.0081,
  "Varanasi", 25.3176, 82.9739,
  "Riga", 56.9496, 24.1052,
  "Paris", 48.8566, 2.3522,
  "Lyon", 45.7640, 4.8357,
  "Brussels", 50.8476, 4.3572,
  "Amsterdam", 52.3676, 4.9041,
  "Frankfurt", 50.1109, 8.6821,
  "Munich", 48.1351, 11.5820,
  "Marseille", 43.2965, 5.3698,
  "Rome", 41.9028, 12.4964,
  "Venice", 45.4408, 12.3155,
  "Berlin", 52.5200, 13.4050,
  "Prague", 50.0755, 14.4378,
  "Warsaw", 52.2297, 21.0122,
  "Vilnius", 54.6872, 25.2797,
  "Tallinn", 59.4370, 24.7536,
  "Stockholm", 59.3293, 18.0686,
  "Oslo", 59.9139, 10.7522,
  "Bergen", 60.3913, 5.3221,
  "Knoxville", 35.9606, -83.9207,
  "New Orleans", 29.9511, -90.0715,
  "Dallas", 32.7767, -96.7970,
  "Miami", 25.7617, -80.1918,
  "Asheville", 35.5951, -82.5515,
  "Damascus", 36.6337, -81.7837,
  "Vancouver", 45.6280, -122.6739
)
#places
```

```{r, echo = FALSE, warning=FALSE, message=FALSE}
my_map <- places %>%
  leaflet(width = "100%") %>%
  addTiles() %>%
  addMarkers(popup = places$name) %>% 
  addProviderTiles(providers$Esri.NatGeoWorldMap)

my_map
```

# Hikes

These are only the hikes in US. My hikes in India and Europe aren't included.

| Name                                                                                                                                        | Location           | Notes                                                      |
|---------------------------------------------------------------------------------------------------------------------------------------------|--------------------|------------------------------------------------------------|
| [Alum Cave](https://hikinginthesmokys.com/alum-cave-trail/), [Mount Leconte](https://hikinginthesmokys.com/mt-leconte-via-alum-cave-trail/) | Tennessee, US      | Amazing views; easy hike; didn't go to the top             |
| [Virginia Creeper Trail](https://visitabingdonvirginia.com/landmarks/virginia-creeper-trail)                                                | Virginia, US       | Bike trail; you'll get dirty, so prepare accordingly       |
| [Midnight Hole](https://blueridgemountainlife.com/midnight-hole/)                                                                           | North Carolina, US | Easy trail; small cliffs to dive as well                   |
| [Ozone Falls](https://www.tn.gov/environment/program-areas/na-natural-areas/natural-areas-east-region/east-region-/na-na-ozone-falls.html)  | Tennessee, US      | One of its kind waterfall; you can go behind the waterfall |
| [Cade's Cove](https://www.nps.gov/grsm/planyourvisit/cadescove.htm)                                                                         | Tennessee, US      | Famous for wildlife; I didn't see any 🫠                   |
| [Charlie's Bunion](https://hikinginthesmokys.com/charlies-bunion/)                                                                          | Tennessee, US      | Delivered what's promised; loved it                        |
| [Shuckstack Fire Tower](https://hikinginthesmokys.com/shuckstack-fire-tower/)                                                               | North Carolina, US | Uphill is tougher but the views from tower are worth it.   |
| [Lacamas Creek Park](https://www.alltrails.com/parks/us/washington/lacamas-creek-park)                                                      | Vancouver, WA      | Easy hike with beautiful sceneries                         |

## Bucket List

-   [Big south fork](https://www.nps.gov/biso/index.htm)
-   [Bald falls](https://www.tnvacation.com/local/tellico-plains-bald-river-falls)
-   Max Patch, AT
-   Stone Mountain
-   Rock city Chattanooga
-   Sound of music
-   Falls creek falls
-   Grayson Highlands
-   Cades Cove --- biking

# Adventure Sports

| Sports          | Location                                     | Review                                                                                        |
|-----------------|----------------------------------------------|-----------------------------------------------------------------------------------------------|
| Skydiving       | The Netherlands                              | best adrenaline rush ever                                                                     |
| Rafting         | Jawahar Ghat, Jharkhand Manali, HP           | too hyped                                                                                     |
| Kayaking        | Jawahar Ghat, Jharkhand Ijams, Knoxville, TN | relaxing if water is calm; else tiring but fun                                                |
| Para-sailing    | Jawahar Ghat, Jharkhand                      | fun; kind of like roller coaster                                                              |
| Para-gliding    | Manali, HP                                   | runway to start and landing are best parts                                                    |
| Skiing          | Latvia                                       | easy to learn; difficult to master                                                            |
| Surfing         | Gokarna, Karnataka                           | nothing like it; waves hit you hard                                                           |
| Scuba-diving    | Fort Dickerson Quarry, Knoxville             | not great unless you have clear water                                                         |
| Mountain Biking | Baker Creek Trail, Knoxville                 | amazing trails (easy, medium and hard); artificial ramps for skating and biking are great too |

## Bucket List

-   Cliff Jumping

-   Bungee Jumping
