---
title: Adrenaline Activities
subtitle: Places travelled, hikes, bike rides and adventure sports
summary: Places travelled, hikes, bike rides and adventure sports
author: Harshvardhan
date: '2022-04-20'
slug: my-activities
categories:
  - life
  - thoughts
tags: []
---

# 🧳 Map of Cities Touched

```{r, include = FALSE}
library(leaflet)
library(tidyverse)
places = tribble(
  ~name, ~lat, ~lng,
  "Jhumri Tilaiya", 24.4315, 85.5310,
  "Indore", 22.7196, 75.8577,
  "Ranchi", 23.3441, 85.3096,
  "Dhanbad", 23.7957, 86.4304,
  "Kolkata", 22.5726, 88.3639,
  "New Delhi", 28.6139, 77.2090,
  "Amritsar", 31.6340, 74.8723,
  "Puri", 19.8135, 85.8312,
  "Goa", 15.4909, 73.8278,
  "Gokarna", 14.5479, 74.3188,
  "Mumbai", 19.0760, 72.8777,
  "Udaipur", 24.5854, 73.7125,
  "Ujjain", 23.1765, 75.7885,
  "Agra", 27.1767, 78.0081,
  "Varanasi", 25.3176, 82.9739,
  "Riga", 56.9496, 24.1052,
  "Paris", 48.8566, 2.3522,
  "Lyon", 45.7640, 4.8357,
  "Brussels", 50.8476, 4.3572,
  "Amsterdam", 52.3676, 4.9041,
  "Frankfurt", 50.1109, 8.6821,
  "Munich", 48.1351, 11.5820,
  "Marseille", 43.2965, 5.3698,
  "Rome", 41.9028, 12.4964,
  "Venice", 45.4408, 12.3155,
  "Berlin", 52.5200, 13.4050,
  "Prague", 50.0755, 14.4378,
  "Warsaw", 52.2297, 21.0122,
  "Vilnius", 54.6872, 25.2797,
  "Tallinn", 59.4370, 24.7536,
  "Stockholm", 59.3293, 18.0686,
  "Oslo", 59.9139, 10.7522,
  "Bergen", 60.3913, 5.3221,
  "Knoxville", 35.9606, -83.9207,
  "New Orleans", 29.9511, -90.0715,
  "Dallas", 32.7767, -96.7970,
  "Miami", 25.7617, -80.1918,
  "Asheville", 35.5951, -82.5515,
  "Damascus", 36.6337, -81.7837,
  "Vancouver", 45.6280, -122.6739,
  "Portland", 45.5152, -122.6784
)
#places
```

```{r, echo = FALSE, warning=FALSE, message=FALSE}
my_map = places %>%
  leaflet(width = "100%") %>%
  addTiles() %>%
  addMarkers(popup = places$name) %>% 
  addProviderTiles(providers$Esri.NatGeoWorldMap)

my_map
```

# ⛰️ Hikes

These are only the hikes in US. My hikes in India and Europe aren't included.

+---------------------------------------------------------------------------------------------------------------------------------------------+---------------------------+--------------------------------------------------------------+
| Name                                                                                                                                        | Location                  | Notes                                                        |
+=============================================================================================================================================+===========================+==============================================================+
| [Alum Cave](https://hikinginthesmokys.com/alum-cave-trail/), [Mount Leconte](https://hikinginthesmokys.com/mt-leconte-via-alum-cave-trail/) | Tennessee, US             | Amazing views; easy hike; didn't go to the top               |
+---------------------------------------------------------------------------------------------------------------------------------------------+---------------------------+--------------------------------------------------------------+
| [Virginia Creeper Trail](https://visitabingdonvirginia.com/landmarks/virginia-creeper-trail)                                                | Virginia, US              | Bike trail; you'll get dirty, so prepare accordingly         |
+---------------------------------------------------------------------------------------------------------------------------------------------+---------------------------+--------------------------------------------------------------+
| [Midnight Hole](https://blueridgemountainlife.com/midnight-hole/)                                                                           | North Carolina, US        | Easy trail; small cliffs to dive as well                     |
+---------------------------------------------------------------------------------------------------------------------------------------------+---------------------------+--------------------------------------------------------------+
| [Ozone Falls](https://www.tn.gov/environment/program-areas/na-natural-areas/natural-areas-east-region/east-region-/na-na-ozone-falls.html)  | Tennessee, US             | One of its kind waterfall; you can go behind the waterfall   |
+---------------------------------------------------------------------------------------------------------------------------------------------+---------------------------+--------------------------------------------------------------+
| [Cade's Cove](https://www.nps.gov/grsm/planyourvisit/cadescove.htm)                                                                         | Tennessee, US             | Famous for wildlife; I didn't see any 🫠                     |
+---------------------------------------------------------------------------------------------------------------------------------------------+---------------------------+--------------------------------------------------------------+
| [Charlie's Bunion](https://hikinginthesmokys.com/charlies-bunion/)                                                                          | Tennessee, US             | Delivered what's promised; loved it                          |
+---------------------------------------------------------------------------------------------------------------------------------------------+---------------------------+--------------------------------------------------------------+
| [Shuckstack Fire Tower](https://hikinginthesmokys.com/shuckstack-fire-tower/)                                                               | North Carolina, US        | Uphill is tougher but the views from tower are worth it.     |
+---------------------------------------------------------------------------------------------------------------------------------------------+---------------------------+--------------------------------------------------------------+
| [Lacamas Creek Park](https://www.alltrails.com/parks/us/washington/lacamas-creek-park)                                                      | Vancouver, WA             | Easy hike with beautiful sceneries                           |
+---------------------------------------------------------------------------------------------------------------------------------------------+---------------------------+--------------------------------------------------------------+
| [Dry Creek Falls](https://www.oregonhikers.org/field_guide/Dry_Creek_Falls_Hike)                                                            | Hood River County, Oregon | Mountainous terrain, fire-parched trees, beautiful waterfall |
+---------------------------------------------------------------------------------------------------------------------------------------------+---------------------------+--------------------------------------------------------------+
| [Clingman's Dome](https://www.nps.gov/grsm/planyourvisit/clingmansdome.htm)                                                                 | North Carolina, US        | Easy walk, beautiful views                                   |
+---------------------------------------------------------------------------------------------------------------------------------------------+---------------------------+--------------------------------------------------------------+
| [Grotto Falls, Trillium Gap Trail](https://www.nps.gov/grsm/planyourvisit/grotto-falls.htm)                                                 | North Carolina, US        | Good hike, very crowded though                               |
+---------------------------------------------------------------------------------------------------------------------------------------------+---------------------------+--------------------------------------------------------------+

### 🪣 Bucket List

-   [Big south fork](https://www.nps.gov/biso/index.htm)
-   [Bald falls](https://www.tnvacation.com/local/tellico-plains-bald-river-falls)
-   Max Patch
-   Stone Mountain
-   Rock city Chattanooga
-   Sound of music
-   Falls creek falls
-   Grayson Highlands
-   Cades Cove --- biking

# 🚴 Bike Rides

Since I got my VanMoof, I've been going on bike trips every so often.

+-----------------------------------------------------------------------------------------------------------------------------------------------+---------------+------------------------------------------------------------------------------------------------------------+
| Name                                                                                                                                          | Location      | Notes                                                                                                      |
+===============================================================================================================================================+===============+============================================================================================================+
| [Baker Creek Preserve](https://www.alltrails.com/trail/us/tennessee/sycamore-and-floyd-fox-loop)                                              | Knoxville, TN | [First](https://www.instagram.com/reel/CdMcHPNFsQAN9b73chMOxVrICo5G8Hi97AUlLE0/) mountain bike experience. |
+-----------------------------------------------------------------------------------------------------------------------------------------------+---------------+------------------------------------------------------------------------------------------------------------+
| [Burnt Bridge Creek Greenway](https://www.alltrails.com/trail/us/washington/burnt-bridge-creek-greenway-trail)                                | Vancouver, WA | Easy and simple ride.                                                                                      |
+-----------------------------------------------------------------------------------------------------------------------------------------------+---------------+------------------------------------------------------------------------------------------------------------+
| [Salmon Creek Trail](https://www.alltrails.com/trail/us/washington/salmon-creek-trail)                                                        | Vancouver, WA | Great scenery, spotted a deer.                                                                             |
+-----------------------------------------------------------------------------------------------------------------------------------------------+---------------+------------------------------------------------------------------------------------------------------------+
| [Lewis-Clark Discovery Greenway](https://www.alltrails.com/trail/us/washington/lewis-clark-discovery-greenway-to-renaissance-trail)           | Vancouver, WA | Passes through some cool architecture.                                                                     |
+-----------------------------------------------------------------------------------------------------------------------------------------------+---------------+------------------------------------------------------------------------------------------------------------+
| [Waterfront Loop](https://www.portland.gov/transportation/walking-biking-transit-safety/suggested-rides#toc-the-classic-waterfront-loop-ride) | Portland, OR  | 11/10.[^1]                                                                                                 |
+-----------------------------------------------------------------------------------------------------------------------------------------------+---------------+------------------------------------------------------------------------------------------------------------+
| [Third Creek Greenway](https://www.alltrails.com/trail/us/tennessee/third-creek-greenway)                                                     | Knoxville, TN | Easy, close to UT                                                                                          |
+-----------------------------------------------------------------------------------------------------------------------------------------------+---------------+------------------------------------------------------------------------------------------------------------+
| [Neyland Greenway, Cheerokee Farmway](https://goo.gl/maps/tnrEPirDPon88hBu8)                                                                  | Knoxville, TN | Absolutely beautiful; never crowded.                                                                       |
+-----------------------------------------------------------------------------------------------------------------------------------------------+---------------+------------------------------------------------------------------------------------------------------------+

[^1]: Bonus point for [Dea's](https://deabardhoshi.medium.com) company.

# 🪂Adventure Sports

+------------------------------------+-------------------------------------+-----------------------------------------------------------------------------------------------+
| Sports                             | Location                            | Review                                                                                        |
+====================================+=====================================+===============================================================================================+
| Skydiving                          | The Netherlands                     | best adrenaline rush ever                                                                     |
+------------------------------------+-------------------------------------+-----------------------------------------------------------------------------------------------+
| Rafting (still-water, white-water) | Jawahar Ghat, Jharkhand             | too hyped; waves make things hard but also more fun                                           |
|                                    |                                     |                                                                                               |
|                                    | Manali, HP                          |                                                                                               |
+------------------------------------+-------------------------------------+-----------------------------------------------------------------------------------------------+
| Kayaking                           | Jawahar Ghat, Jharkhand             | relaxing if water is calm; else tiring but fun                                                |
|                                    |                                     |                                                                                               |
|                                    | Ijams, Knoxville, TN                |                                                                                               |
|                                    |                                     |                                                                                               |
|                                    | Lake Merwin, WA                     |                                                                                               |
+------------------------------------+-------------------------------------+-----------------------------------------------------------------------------------------------+
| Para-sailing                       | Jawahar Ghat, Jharkhand             | fun; kind of like roller coaster                                                              |
+------------------------------------+-------------------------------------+-----------------------------------------------------------------------------------------------+
| Para-gliding                       | Manali, HP                          | runway to start and landing are best parts                                                    |
+------------------------------------+-------------------------------------+-----------------------------------------------------------------------------------------------+
| Skiing                             | Latvia                              | easy to learn; difficult to master                                                            |
+------------------------------------+-------------------------------------+-----------------------------------------------------------------------------------------------+
| Surfing                            | Gokarna, Karnataka                  | nothing like it; waves hit you hard                                                           |
+------------------------------------+-------------------------------------+-----------------------------------------------------------------------------------------------+
| Scuba-diving                       | Fort Dickerson Quarry, Knoxville    | not great unless you have clear water                                                         |
+------------------------------------+-------------------------------------+-----------------------------------------------------------------------------------------------+
| Mountain Biking                    | Baker Creek Trail, Knoxville, TN    | amazing trails (easy, medium and hard); artificial ramps for skating and biking are great too |
+------------------------------------+-------------------------------------+-----------------------------------------------------------------------------------------------+
| Bungee Jumping                     | Pacific Northwest Bridge, Amboy, WA | jump is scary; everything else is fun! the high only lasts for a few seconds though           |
+------------------------------------+-------------------------------------+-----------------------------------------------------------------------------------------------+
| Paddle-boarding                    | Ijams, Knoxville, TN                | use paddle board for diving; paddle-boarding is boring by itself :)                           |
+------------------------------------+-------------------------------------+-----------------------------------------------------------------------------------------------+

### 🪣 Bucket List

-   Cliff Jumping

-   ~~Bungee Jumping~~
