---
title: Elon Gets Twitter
subtitle: 'What else can you say in less than 280 characters'
summary: 'What else can you say in less than 280 characters'
author: Harshvardhan
date: '2022-05-08'
slug: elon-gets-twitter
categories:
  - life
  - thoughts
tags: []
draft: true
---

[Opinion \| Elon Musk Got Twitter Because He Gets Twitter - The New York Times](https://www.nytimes.com/2022/04/27/opinion/elon-musk-twitter.html)

> The man loves Twitter. He tweets as if he had been raised by the blue bird and the [fail whale](https://archive.ph/o/343qJ/https://www.techopedia.com/definition/1987/fail-whale).

> Jack Dorsey, a Twitter co-founder and former chief executive, always wanted it to be something else. Something it wasn't, and couldn't be. "The purpose of Twitter is to serve the public conversation," he [said](https://archive.ph/o/343qJ/https://www.intelligence.senate.gov/sites/default/files/documents/os-jdorsey-090518.pdf) in 2018. Twitter began "[measuring conversational health](https://archive.ph/o/343qJ/https://blog.twitter.com/en_us/topics/company/2018/measuring_healthy_conversation)" and trying to tweak the platform to burnish it. Sincere as the effort was, it was like those liquor ads advising moderation. You don't get people to drink less by selling them whiskey.

[The Real History of Twitter, in Brief](https://www.lifewire.com/history-of-twitter-3288854)

> Twitter began as an idea that Twitter co-founder Jack Dorsey ([\@Jack](https://twitter.com/jack)) had in 2006. Dorsey had originally imagined Twitter as an SMS-based communications platform. Groups of friends could keep tabs on what each other were doing based on their status updates. Like texting, but not.

```{r echo=FALSE, eval=FALSE}
blogdown::shortcode('tweet', '1517215066550116354')
blogdown::shortcode('tweet', "1519469891455234048")
blogdown::shortcode('tweet', "1519415674111672325")
blogdown::shortcode('tweet', "1517215066550116354")
blogdown::shortcode('tweet', "1507259709224632344")
blogdown::shortcode('tweet', "1507041396242407424")
blogdown::shortcode('tweet', "1512785529712123906")
blogdown::shortcode('tweet', "1507777261654605828")
blogdown::shortcode('tweet', "1519179787163652099")
blogdown::shortcode('tweet', "1519020176884305920")
blogdown::shortcode('tweet', "1470858546153762819")
blogdown::shortcode('tweet', "1517707521343082496")
blogdown::shortcode('tweet', "1518772756069773313")
blogdown::shortcode('tweet', "1521523506512470018")
blogdown::shortcode('tweet', "1511322655609303043")
```

# This one blows my mind.

It will go into textbooks for corporate communication failure, at least at business schools.

```{r, echo = FALSE, eval=FALSE}
library(rtweet)

last_tweet_id <- "1526250477456965634"

lookup_thread <- function(status_id, tweets = NULL) {
  tweet <- rtweet::lookup_tweets(status_id)
  tweets <- dplyr::bind_rows(tweet, tweets)
  if (is.na(tweet$reply_to_status_id)) {
    return(tweets)
  } else {
    lookup_thread(tweet$reply_to_status_id, tweets)
  }
}

# gather from last tweet to first tweet in the thread,
# following the trail of reply_to_status_id
thread <- lookup_thread(last_tweet_id)

thread <- paste0("https://twitter.com/", thread$screen_name, "/status/", 
       thread$status_id)

htmltools::tagList(
  lapply(thread, tweetrmd::tweet_embed, plain = TRUE)
)
```
