---
title: Notes on R
subtitle: Functions et al
summary: Functions et al
author: Harshvardhan
date: '2022-02-03'
slug: some-r-functions
categories:
  - R
tags: []
---

On one fine day when I have enough time, they'll all be wrapped into a package hosted on my Github. Until then, this page in their home.

```{r}
library(tidyverse)
```

## Show in Excel

```{r}
show_in_excel = function(.data)
{
   temp = paste0(tempfile(), ".csv")
   write.csv(.data, temp)
   fs::file_show(path = temp)
}
```

This can be used with pipes too.

```{r eval=FALSE}
iris %>% 
  show_in_excel()
```

## Convert Missing Values to Zero

This function converts missing values in a vector to zero.

```{r}
n2z = function(x)
{
   x = ifelse(is.na(x), 0, x)
   return(x)
}
```

## Finding Index of Missing Elements

This function returns index of elements which are missing. Very useful in finding what observations are missing.

```{r}
which.na = function(x)
{
   return(which(is.na(x)))
}
```

## Removing Rows Based on Missing Values in a Column

Sometimes, I do not want to `na.omit()` because it will treat all features equally. I want to check values only for one column.

```{r}
na.rm.feature = function(x, colname)
{
   nas = which(is.na(x[,colname]))
   x = x[-nas, ]
   return(x)
}
```

## Find row where a condition is satisfied

This function can find observations that satisfy a condition. Typically, they are useful in finding specific elements. It is kind of a wrapper around dplyr's `filter()`.

```{r}
which.this = function(df, x)
{
  df %>% 
    filter(eval(parse(text = x)))
}
```

#### Example

```{r}
which.this(iris, "Sepal.Length > 6.5") %>% 
  head()
```

## Remove commas, dollars, or any other such characters

The code below replaces all commas with nothing.

```{r}
x = "300,000"
x = gsub(",", "", x)
print(x)
```

## GGPlot2 Theme

See [official guide](https://ggplot2.tidyverse.org/reference/theme.html) for more details. Also see [Benjamin's blog](https://benjaminlouis-stat.fr/en/blog/2020-05-21-astuces-ggplot-rmarkdown/).

The default plot looks like this.

```{r}
iris %>% 
  ggplot(aes(x = Sepal.Length, y = Sepal.Width, colour = Species)) +
  geom_point()
```

Once I run and set my theme, its way prettier.

```{r}
theme_h = function(base_size = 14) {
  theme_bw(base_size = base_size) %+replace%
    theme(
      # Specify plot title
      plot.title = element_text(size = rel(1), face = "bold", margin = margin(0,0,5,0), hjust = 0),
      # Specifying grid and border
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      # Specidy axis details
      axis.title = element_text(size = rel(0.85), face = "bold", family="serif"),
      axis.text = element_text(size = rel(0.70), family="serif"),
      axis.line = element_line(color = "black", arrow = arrow(length = unit(0.3, "lines"), type = "closed")),
      # Specify legend details
      legend.title = element_text(size = rel(0.85), face = "bold"),
      legend.text = element_text(size = rel(0.70), face = "bold", family="serif"),
      legend.key = element_rect(fill = "transparent", colour = NA),
      legend.key.size = unit(1.5, "lines"),
      legend.background = element_rect(fill = "transparent", colour = NA),
      # Remove default background
      strip.background = element_rect(fill = "#17252D", color = "#17252D"),
      strip.text = element_text(size = rel(0.85), face = "bold", color = "white", margin = margin(5,0,5,0))
    )
}

theme_set(theme_h())
```

```{r}
iris %>% 
  ggplot(aes(x = Sepal.Length, y = Sepal.Width, colour = Species)) +
  geom_point()
```

I like the arrowed axes and serif fonts.

## Better Quality Images in R Markdown

Using `.svg` as the image output format gives much better graphics quality than the default option. To use that, include the following code in R Markdown. [Source](https://stackoverflow.com/questions/64602680/how-to-set-svg-as-default-rendering-in-rmarkdown).

```{r, eval = FALSE}
knitr::opts_chunk$set(dev = 'svg') # set output device to svg
```
