---
title: Street Maps (of Some Cities)
subtitle: Some beautiful street maps
summary: "Over the last few days, I dabbled with maps in R. Two days ago, I made a map of all the cities I've visited. Today, I thought to make street maps of some of them (and other cool cities)."
author: Harshvardhan
date: '2022-05-28'
slug: city-maps
categories:
  - R
tags: []
---

Over the last few days, I dabbled with maps in R. Two days ago, I made a map of all the cities I've visited. Today, I thought to make street maps of some of them (and other cool cities).

The dark grey lines are highways and roadways, light grey lines are other streets and blue is water.

# Jhumri Tilaiya, India

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(osmdata)

city_mapper = function(city)
{
   lts = getbb(city)
   
   streets = getbb(city)%>%
      opq()%>%
      add_osm_feature(key = "highway", 
                      value = c("motorway", "primary", 
                                "secondary", "tertiary")) %>%
      osmdata_sf()
   
   small_streets = getbb(city)%>%
      opq()%>%
      add_osm_feature(key = "highway", 
                      value = c("residential", "living_street",
                                "unclassified",
                                "service", "footway")) %>%
      osmdata_sf()
   
   river = getbb(city)%>%
      opq()%>%
      add_osm_feature(key = "waterway", value = "river") %>%
      osmdata_sf()
   
   p = ggplot() +
      geom_sf(data = streets$osm_lines,
              inherit.aes = FALSE,
              color = "#282828", #3C280D
              size = .5,
              alpha = .7) +
      geom_sf(data = small_streets$osm_lines,
              inherit.aes = FALSE,
              color = "#909090", #795C34
              size = .4,
              alpha = .4) +
      geom_sf(data = river$osm_lines,
              inherit.aes = FALSE,
              color = "#03026F",
              size = .7,
              alpha = .8) +
      coord_sf(xlim = c(lts[1], lts[3]), 
               ylim = c(lts[2], lts[4]),
               expand = FALSE) +
      theme_void() +
      labs(caption = "Learn more: harsh17.in/city-maps")
   
   return(p)
   
}
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
home = city_mapper("Jhumri Tilaiya India")
home
```

# Indore, India

```{r, echo=FALSE, warning=FALSE, message=FALSE}
indore = city_mapper("Indore India")
indore
```

# Riga, Latvia

```{r, echo=FALSE, warning=FALSE, message=FALSE}
riga = city_mapper("Riga Latvia")
riga
```

# Knoxville, United States

```{r, echo=FALSE, warning=FALSE, message=FALSE}
knoxville = city_mapper("Knoxville United States")
knoxville
```

# New Delhi, India

```{r, echo=FALSE, warning=FALSE, message=FALSE}
delhi = city_mapper("New Delhi India")
delhi
```

# Sydney, Australia

```{r, echo=FALSE, warning=FALSE, message=FALSE}
sydney = city_mapper("Sydney Australia")
sydney
```

# New York, United States

```{r, echo=FALSE, warning=FALSE, message=FALSE}
newyork = city_mapper("New York United States")
newyork
```

# Boston, United States

```{r, echo=FALSE, warning=FALSE, message=FALSE}
boston = city_mapper("Boston United States")
boston
```

# London, United Kingdom

```{r, echo=FALSE, warning=FALSE, message=FALSE}
london = city_mapper("London United Kingdom")
london
```

------------------------------------------------------------------------

The function to generate these is not complicated. For a detailed tutorial, [see this tutorial](https://ggplot2tutor.com/tutorials/streetmaps?utm_source=pocket_mylist). Here's the function that I used.

```{r eval = FALSE}
library(tidyverse)
library(osmdata)

city_mapper = function(city)
{
   lts = getbb(city)
   
   streets = getbb(city)%>%
      opq()%>%
      add_osm_feature(key = "highway", 
                      value = c("motorway", "primary", 
                                "secondary", "tertiary")) %>%
      osmdata_sf()
   
   small_streets = getbb(city)%>%
      opq()%>%
      add_osm_feature(key = "highway", 
                      value = c("residential", "living_street",
                                "unclassified",
                                "service", "footway")) %>%
      osmdata_sf()
   
   river = getbb(city)%>%
      opq()%>%
      add_osm_feature(key = "waterway", value = "river") %>%
      osmdata_sf()
   
   p = ggplot() +
      geom_sf(data = streets$osm_lines,
              inherit.aes = FALSE,
              color = "#282828", #3C280D
              size = .5,
              alpha = .7) +
      geom_sf(data = small_streets$osm_lines,
              inherit.aes = FALSE,
              color = "#909090", #795C34
              size = .4,
              alpha = .4) +
      geom_sf(data = river$osm_lines,
              inherit.aes = FALSE,
              color = "#03026F",
              size = .7,
              alpha = .8) +
      coord_sf(xlim = c(lts[1], lts[3]), 
               ylim = c(lts[2], lts[4]),
               expand = FALSE) +
      theme_void() +
      labs(caption = "Learn more: harsh17.in/city-maps")
   
   return(p)
   
}
```
