---
title: Notes on Python
subtitle: Some things that I'll likely forget
summary: Some things that I'll likely forget
author: Harshvardhan
date: '2022-02-08'
slug: notes-on-python
categories:
  - Python
tags: []
---

# Installing Packages

There are ~~two~~ three ways: `pip`, `conda` and within Jupyter Hub. If you are using Anaconda (Navigator), use `conda`. `pip` will install for all environments; `conda` will install for conda environment only. Put either of these codes in Terminal.

```{r eval = FALSE}
# using pip
python -m pip install <package>

# using conda
conda install <package>
```

If it is a one off use case (which it usually never is), use the following command.

```{r eval = FALSE}
!pip install <package>
```

# Methods to Print

There are two methods: `print()` and `pprint()`. They both serve the save cause but `pprint()` is better at displaying complex data structures such as list of lists or JSON files.

# Adding New Rows to Data Frame

It is more efficient to create a list of rows first and then convert it to a pandas data frame. As qmeeus [said](https://stackoverflow.com/questions/55967976/python-panda-append-dataframe-in-loop?noredirect=1&lq=1) on SO,

> Pandas dataframes do not work as a list, they are much more complex data structures and appending is not really considered the best approach.

```{r eval = FALSE}
df = pd.DataFrame()
r_list = []
for r in range(result_list):
  results = r_list.append(result_list)
  
df = pd.concat([results, df], axis=0).reset_index(drop=True)
```
